\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{float}

\title{Replicating Table 1, Panel B from\\Chernozhukov et al.\ (2017)}
\author{Aaron Chotzen-Jenner}
\date{\today}

\begin{document}
\maketitle

\section{The Original Paper}

Chernozhukov, Chetverikov, Demirer, Duflo, Hansen, and Newey (2017),
``Double/Debiased/Neyman Machine Learning of Treatment Effects,''
\textit{American Economic Review: Papers \& Proceedings}, 107(5): 261--265,
introduces the Double/Debiased Machine Learning (Double ML) estimator for
causal inference in settings where high-dimensional nuisance parameters must
be estimated. The central research question is how to obtain valid,
root-$N$ consistent estimates of a treatment effect parameter $\theta_0$
when the confounding functions are estimated using flexible machine learning
methods that introduce regularization bias. In simpler terms, the paper
explains how to measure the true effect of a treatment when machine learning
is used to control for other factors that may influence the outcome.

The authors approach this problem with two main methods. First, the
estimating equation is made Neyman-orthogonal, meaning it is locally
insensitive to small errors in the nuisance parameter estimates. In other
words, the equation is designed in a way that small mistakes in the machine
learning predictions have very little impact on the final treatment effect
estimate. Second, cross-fitting is used to split the sample into folds,
training the nuisance functions (everything else that needs to be estimated
besides the main effect) on held-out data to avoid overfitting bias.
Together, these two steps remove the regularization bias and therefore not
distorted by error from the machine learning models.

\section{The Result of Interest}

The specific result I replicate is the Mean Average Treatment Effect (ATE)
reported in Table 1, Panel B, row ``ATE (2 fold),'' Ensemble column of the
online appendix. This result corresponds to the partially linear model with
2-fold cross-fitting. The paper reports a Mean ATE of \$9,043 with a
standard error of 1,432.

This result is the headline finding of the paper\@: after flexibly controlling
for income and other confounding factors, 401(k) eligibility increases net
financial assets by roughly \$9,000. The underlying model estimates the
relationship between financial assets, 401(k) eligibility, and a set of
demographic and economic control variables. Rather than relying on a single
machine learning algorithm, the Ensemble method combines predictions from
several approaches, including Lasso, Random Forests, Boosting, and Neural
Networks. It assigns more weight to the methods that perform best
out-of-sample, as the goal is to obtain a more reliable estimate than any
single method alone.

I chose this result to replicate because it is the paper's primary empirical
finding, corresponds to the default specification in the replication code
(\texttt{Master.R}), and illustrates the central idea of Double ML, which is
using machine learning to control for complex confounding while still
producing a credible estimate of a causal treatment effect.

\section{Data and Methods}

The data come from the 1991 Survey of Income and Program Participation
(SIPP), the same dataset used in Chernozhukov and Hansen (2004). The sample
contains 9,915 observations. The outcome variable \texttt{net\_tfa} is net
total financial assets, measured in dollars. The treatment variable
\texttt{e401} is a binary indicator for eligibility to enroll in a 401(k)
plan. The nine covariates used in the model are age, income, years of
education, family size, and binary indicators for married status,
two-earner household, defined benefit pension, IRA participation, and
home ownership.

The partially linear model estimated here takes the form
\begin{align}
    Y &= \theta_0 D + g_0(Z) + \zeta, \quad E[\zeta \mid Z, D] = 0,
        \label{eq:outcome} \\
    D &= m_0(Z) + \nu, \quad E[\nu \mid Z] = 0, \label{eq:treatment}
\end{align}
where $Y$ is net financial assets, $D$ is 401(k) eligibility, $Z$ is the
vector of covariates, $\theta_0$ is the Average Treatment Effect (ATE) of
interest, and $g_0(Z)$ and $m_0(Z)$ are unknown nuisance functions estimated
via machine learning. Equation~(\ref{eq:outcome}) is the outcome equation,
where $\theta_0$ enters linearly and $g_0(Z)$ captures the flexible effect of
controls on assets. Equation~(\ref{eq:treatment}) is the treatment equation,
where $m_0(Z) := E[D \mid Z]$ is the propensity score --- the probability of
401(k) eligibility given observables.

The key insight of Double ML is that plugging ML estimates of $g_0$ and $m_0$
directly into a naive regression for $\theta_0$ introduces regularization bias.
Double ML addresses this problem with two key ideas. First, the estimating
equation is made Neyman-orthogonal using an orthogonal score such as
\begin{equation}
    \psi(W,\theta,\eta) := \bigl(Y - \hat{g}(Z) - \theta(D -
        \hat{m}(Z))\bigr)(D - \hat{m}(Z)), \label{eq:score}
\end{equation}
which is locally insensitive to small errors in the nuisance estimates
$\hat{g}$ and $\hat{m}$. This amounts to a residuals-on-residuals
regression: the outcome residual $Y - \hat{g}(Z)$ is regressed on the
treatment residual $D - \hat{m}(Z)$, removing the influence of covariates
before estimating the treatment effect. Second, cross-fitting is used to
avoid overfitting: the nuisance functions are trained on a held-out fold
and applied to the remaining fold, so that the same observations are never
used for both nuisance estimation and moment evaluation.

\texttt{preprocess.R} loads the raw Stata file \texttt{sipp1991.dta} from
\texttt{input/}, selects the eleven relevant variables, and writes a clean
CSV to \texttt{temp/clean\_data.csv}. \texttt{analysis.R} loads that file,
sources the original helper functions \texttt{ML\_Functions.R} and
\texttt{Moment\_Functions.R} from \texttt{input/}, and calls
\texttt{DoubleML()} with the partially linear specification above and 2-fold
cross-fitting. To reduce computation time, I repeated the estimation over
10 random sample splits rather than the 100 used in the original paper.
I used parallel computation across 4 cores of my computer to accelerate
estimation. The Mean ATE, Median ATE, and their standard errors are computed
following the paper's formulas exactly. Results are written to
\texttt{output/tables/main\_result.tex}, which is included directly in this
document via \texttt{\textbackslash input\{\}}.

\section{Replication Results}

\input{../output/tables/main_result.tex}

\input{../output/tables/comparison_table.tex}

\noindent This replication closely reproduces the headline result. The
Ensemble estimate of the ATE is \$8,970, compared to \$9,043 reported in
the paper, a difference of \$73, or less than one percent. The
replication standard error of 1,407 is nearly identical to the
paper's 1,432, and the 95\% confidence interval [\$6,212, \$11,728]
comfortably contains the published point estimate. The Median ATE of
\$9,017 is also close to the paper's median of \$9,061, confirming that the
result is not driven by outlier splits.

These small discrepancies are expected for a replication like this and can
be easily explained. The original paper averages over 100 random sample
splits, but this replication uses only 10. With fewer splits, individual
random partitions carry more weight, so the mean estimate is more sensitive
to any one particularly favorable or unfavorable draw of the folds. The
fact that the two estimates still agree to within one percent despite this
reduction suggests the procedure is very stable. This is consistent with
the Neyman orthogonality framework, which suggests that the final treatment
effect estimate should not be overly sensitive to small errors in the
nuisance estimates or to a particular sample split.

Table 2 shows the method-by-method comparison. The overall pattern is
also consistent between the replication and the paper. RLasso produces the
lowest ATE in both (\$7,603 vs.\ \$7,718), reflecting the tendency of
$\ell_1$-penalized linear models to underfit when extrapolation is required
across folds. Random Forest exhibits the largest difference (\$8,757
vs.\ \$9,180), which may reflect the greater sensitivity of tree-based
methods to the particular sample splits used in estimation. Random Forest
and Neural Net sit at the higher end in both cases, and the Ensemble sits
close to them, as expected given that it up-weights whichever methods
minimize out-of-sample prediction error. Standard errors are broadly
comparable across all six methods, with differences generally well below
100. The consistency of these results across very different machine
learning approaches can also be seen as a substantive finding, confirming
the paper's central claim that Neyman orthogonality makes the treatment
effect estimate robust to the choice of nuisance estimator.

\section{What I Learned}

\noindent Once the helper functions were sourced correctly, reproducing the
estimation was fairly straightforward. The hardest part was path management,
as the original code assumed all files lived in the same working directory,
while the Gentzkow-Shapiro structure separates \texttt{input/},
\texttt{code/}, and \texttt{temp/}. In particular,
\texttt{Moment\_Functions.R} internally sources \texttt{ML\_Functions.R}
with a bare filename, which required patching to use an absolute path. This
was a pure reproducibility issue\@: code that works perfectly in the author's
environment can break when placed in a different directory structure. This
happened to me before I made the fix. If I were the original authors, the
one thing that would have made replication easier is absolute rather than
relative paths in the helper files, along with a brief README note
specifying how the scripts should be run.

More broadly, this replication taught me that Double ML is amazingly stable.
Six very different machine learning methods (RLasso, Trees, Random Forest,
Boosting, Neural Net, and Ensemble) all produce estimates between
approximately \$7,900 and \$9,200, with substantial overlap in their
confidence intervals. This stability is exactly what the Neyman orthogonality
theory predicts, and seeing it empirically with real-world data was something
I could have learned only from doing this replication.

\begin{thebibliography}{9}
\bibitem{paper}
Chernozhukov, V., Chetverikov, D., Demirer, M., Duflo, E., Hansen, C.,
and Newey, W. (2017). ``Double/Debiased/Neyman Machine Learning of
Treatment Effects.'' \textit{American Economic Review: Papers \&
Proceedings}, 107(5): 261--265.
\end{thebibliography}

\end{document}
